\vspace*{-8mm}\textbf{\textsc{Novelty Detection for Semantic Place Categorization}}

### 2011-06-23 thesis/summary-pt.tex
Added portuguese sumary version and scons rules.

%% English version: comment first, uncomment second
%\usepackage[portuguese]{babel}  % Portuguese
\usepackage[english]{babel}     % English
\usepackage{graphicx}           % images .png or .pdf w/ pdflatex OR .eps w/ latex
\usepackage{times}              % use Times type-1 fonts
\usepackage[utf8]{inputenc}     % 8 bits using UTF-8
\usepackage{url}                % URLs
\usepackage{multicol}           % twocolumn, etc
\usepackage{float}              % improve figures & tables floating
\usepackage[tableposition=top]{caption} % captions
%% English version: comment first (maybe)
%\usepackage{indentfirst}        % portuguese standard for paragraphs
%\usepackage{parskip}

@@

\title{\vspace*{-8mm}\textbf{\textsc{Novelty Detection for Semantic Place Categorization}}}
\author{\emph{André Susano Pinto}$^\dag$\\[2mm]
        \small{Supervised by: Andrzej Pronobis$^\ddag$, Luis Paulo Reis$^\dag$}\\
        \small{$^\dag$Faculdade de Engenharia da Universidade do Porto, Portugal}\\
        \small{$^\ddag$Royal Institute of Technology (KTH), Stockholm, Sweden}}
%\url{andresusanopinto@gmail.com}\\
\date{}
\maketitle
%no page number 
\thispagestyle{empty}

@@

\section{Motivation}
For a long time humanity has fantasized that one day robots will walk among us.
They will move and be able to interact with us, understand our concepts and
be able to reason.
An important step on that direction is endowing robots with knowledge of human concepts and
semantics. However its unrealistic to believe on the possibility of extensively
describing human concepts and semantics to robots on a pre-deploy phase, and so, robots must be able
to adapt and learn when confronted with novel situations.
Detection of novel situations, where the robot knowledge is not sufficient, plays then an important
role for the subsequent adapting and learning of new concepts and its this thesis main topic.

Additionally in the context of mobile robotics, spatial concepts and semantics play a main role to perform
complex human-like tasks and human interactions. For handling those, a robot builds a representation
of space extended with semantic properties, process which is known as semantic mapping.
This representation identifies spatial entities and classifies them according to their meaning to
humans allowing the robot to reason at a very high-level.
For example, in a room scale, humans categorize space in kitchens, bedrooms, corridors,
office, theater, computer lab, etc\dots, by relating room areas with those categories, that describing expected
properties and actions that take place on them, a robot can solve tasks like:
``Where are the cornflakes?''

\section{Problem and Goals}
The semantic mapping process plays an important role for allowing agents to reason at an high-level
and is totally defined by the knowledge the agent has on the spatial concepts and semantic categories.
For that reason there is an high interest in creating methods that are able to alert the agent for
gaps on its knowledge.
And with that purpose, this thesis, studies novelty detection techniques and how to apply them to
detect novel room categories during a semantic mapping process.
To this end, the semantic mapping process proposed by \cite{pronobis2011phd} is studied and a method
to detect novel categories using graphical models is suggested.

\section{Semantic Mapping}
Semantic mapping is the process by which an agent builds a representation of the environment and
enhances the identified entities with semantic values (i.e.\ human semantics).
In this thesis the semantic mapping process proposed by \cite{pronobis2011phd} is used as base.

\subsection{Spatial Knowledge Representation}
In order to perform semantic mapping an agent needs to possess knowledge on how to segment and
identify entities existent in space by analysing the sensed data. For that
Pronobis~\cite{pronobis2011semmap} presents a layered knowledge structure:
the \emph{sensory layer} maintains an accurate and exact description of environment in a machine-friendly representation,
the \emph{place layer} employs knowledge on how to segment and identify areas such as rooms,
the \emph{categorical layer} defines knowledge on properties and categories that allow to extract
higher-level information from the sensed data
and the \emph{conceptual layer} maintains knowledge on entities concepts, the relations between them
and information on how to map them to human semantics.

>>>>

%% English version: comment first, uncomment second
\usepackage[portuguese]{babel}  % Portuguese
%\usepackage[english]{babel}     % English
\usepackage{graphicx}           % images .png or .pdf w/ pdflatex OR .eps w/ latex
\usepackage{times}              % use Times type-1 fonts
\usepackage[utf8]{inputenc}     % 8 bits using UTF-8
\usepackage{url}                % URLs
\usepackage{multicol}           % twocolumn, etc
\usepackage{float}              % improve figures & tables floating
\usepackage[tableposition=top]{caption} % captions
%% English version: comment first (maybe)
\usepackage{indentfirst}        % portuguese standard for paragraphs
%\usepackage{parskip}

@@

\title{\vspace*{-8mm}\textbf{\textsc{Novelty Detection for Semantic Place Categorization}}}
\author{\emph{André Susano Pinto}$^\dag$\\[2mm]
        \small{Sob orientação de: Andrzej Pronobis$^\ddag$, Luis Paulo Reis$^\dag$}\\
        \small{$^\dag$Faculdade de Engenharia da Universidade do Porto, Portugal}\\
        \small{$^\ddag$Royal Institute of Technology (KTH), Stockholm, Sweden}}
%\url{andresusanopinto@gmail.com}\\
\date{}
\maketitle
%no page number 
\thispagestyle{empty}

@@

\section{Motivation}
Já à muitos anos que a humanidade sonha que um dia robôs caminharam lado a lado com os humanos.
Acredita-se que eles serão capaz de ser mover e interagir connosco, compreender os nossos conceitos
e serem capazes de pensar.
Um passo importante nesta direcção é a criação de robôs com conhecimento de conceitos e significados
próprios do mundo humano.
No entanto é irrealista acreditar que será possivel descrever e representar em toda a sua extensão
tais conceitos. Assim os agentes deverão ser capazes de aprender e adaptar o seu conhecimento quando
confrontados com novas situações.
A detecção destas novas situações onde o conhecimento de um agente não é suficiente têm um papel importante
na adaptação e aprendizagem de novos conceitos e significados e é por isso o tópico principal desta tese.

Addicionalmente no contêxto de agentes moveis, para facilitar o deslocamento e planeamento, os
agentes constroiem uma representação do ambiente na qual detectam conceitos e propriedades
conceitos espaciais.
Essa representação identifica conceitos e classifica-os de acordo com significados humanos,
permitindo assim ao agente obter uma representação dos conceitos humanos e pensar e planear a um
nível próximo dos mesmos.
Ao processo de construção desta representação é attribuido o nome de \emph{semantic mapping}.
Por exemplo, a uma escala de quartos, os humanos categorizam o espaço em: cozinhas, quartos, corredores,
escritórios, sala de computadores, etc\dots
Ao relacionar quartos com estas categorias que descrevem propriedades e acções percebidas por
humanos, um agente consegue resolver tarefas como: ``Onde estão os ceriais?''

\section{Problema e Objectivos}
Como deliniado na motivação, o processo de \emph{semantic mapping} têm um papel importante na
capacidade pensativa de um agente. Em simultâneo, esse processo está limitado pelo conhecimento
do agente em conceitos e significados do mundo humano. É então importante estudar e analisar
técnicas que sejam capazes de detectar falhas no conhecimento e permitir ao agente agir de maneira
cuidadosa.
É com esse objectivo que a tese aqui descrita se foca no estudo de deteção
de situações novas num processo de \emph{semantic mapping} propôsto por \cite{pronobis2011phd}.
Para esse fim a tese estuda como usar modelos graphicos para detectar novas categorias nas conhecimento
de categorias do agente.

\section{Mapeamento Semântico}
Mapeamento semântico é o processo através do qual um agente constroi uma representação do ambiente
extendida com conceitos e categories com significado especial (i.e.\ conceitos humanos).
Neste tese o processo proposto por \cite{pronobis2011phd} foi usado como base.

\subsection{Conhecimento Espacial}
O conhecimento espacial de agente define todo a informação necessária para o mapeamento semântico,
e inclui informação em como segmentar e identificar entidades e propriedades através dos dados
recolhidos.
Para esse fim Pronobis~\cite{pronobis2011semmap} utiliza uma representação estruturada em camadas:
a \emph{camada de sensores} é responsável por manter uma descrição exacta do ambiente numa representação
de baixo nível;
a \emph{camada de lugares} é responsável por segmentar e identificar areas e conceitos
como quartos;
a \emph{camada de categories} é responsável por definir conhecimento sobre propriadades e categories
e permite ao agente obter informação de mais alto-nível;
e a \emph{camada de conceitos} mantêm conhicimento sobre os diversoes conceitos, relações sobre os mesmos
e informação de como os mapear para conceitos humanos.
